---
title: "Lab 04"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r install-libraries}
library(lubridate)
library(tidyverse)
library(data.table)
```


## Step 1. Read in the data
First download and then read in with data.table:fread()

```{r read-data, cache=TRUE}
if (!file.exists("met_all.gz")){
  download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz", "met_all.gz", method="libcurl", timeout = 60)
}

met <- data.table::fread("met_all.gz")
```
## 2. Prepare the data

Remove temperatures less than -17C and change elev 9999 to missing value code.
```{r remove-low-temp}
met <- met[met$temp >= -17][elev == 9999.0, elev:= NA]
```

Generate a date variable using the functions as.Date() (hint: You will need the following to create a date paste(year, month, day, sep = "-")).

```{r create-ymd, cache=TRUE}
met <- met[, ymd := as.Date(paste(year, month, day, sep = "-"))]
```

Using the data.table::week function, keep the observations of the first week of the month.

```{r}
met[, table(week(ymd))]
met <- met[week(ymd)==31]
```

Compute the mean by station of the variables temp, rh, wind.sp, vis.dist, dew.point, lat, lon, and elev.

```{r check-max}
met[ , .(
  temp      = max(temp),
  rh        = max(rh, na.rm=T), 
  wind.sp   = max(wind.sp, na.rm=T), 
  vis.dist  = max(vis.dist, na.rm=T), 
  dew.point = max(dew.point, na.rm=T), 
  lat       = max(lat, na.rm=T), 
  lon       = max(lon, na.rm=T), 
  elev      = max(elev, na.rm=T)
)]
```
```{r}
met_avg <- met[ , .(
  temp      = mean(temp),
  rh        = mean(rh, na.rm=T), 
  wind.sp   = mean(wind.sp, na.rm=T), 
  vis.dist  = mean(vis.dist, na.rm=T), 
  dew.point = mean(dew.point, na.rm=T), 
  lat       = mean(lat, na.rm=T), 
  lon       = mean(lon, na.rm=T), 
  elev      = mean(elev, na.rm=T)
), by = "USAFID"]

```

Create a region variable for NW, SW, NE, SE based on lon = -98.00 and lat = 39.71 degrees

```{r}
met_avg[, region := fifelse(lon >= -98 & lat > 39.71, "NE", 
                fifelse(lon < -98 & lat > 39.71, "NW",
                fifelse(lon < -98 & lat <= 39.71, "SW",
                "SE")))
    ]
table(met_avg$region)
```

Create a categorical variable for elevation as in the lecture slides

```{r}
met_avg[ ,elev_cat := fifelse(elev > 252, "high", "low")]
```

## 3. Make Viloin plots of dew point temp by design

```{r}
met_avg[!is.na(region)] %>%
  ggplot() +
  geom_violin(mapping = aes( x=1, y= dew.point, color= region, fill=region))+
  facet_wrap(~region, nrow = 1)
```

The highest dew point temperatures are reported inthe southeast.

```{r violin-wind.sp}
met_avg[!is.na(region)& !is.na(wind.sp)] %>%
  ggplot() +
  geom_violin(mapping = aes( x=1, y= wind.sp, color= region, fill=region))+
  facet_wrap(~region, nrow = 1)
```

The highest wind spped seem to occur in the NE.

## 4. Use geom_point with geom_smooth to examine the association between dew point temperature and wind speed by region
Colour points by region
Make sure to deal with NA category
Fit a linear regression line by region
Describe what you observe in the graph

```{r scatterplot-dewpoint-wind.sp}
met_avg[!is.na(region)& !is.na(wind.sp)& !is.na(dew.point)] %>%
  ggplot(mapping = aes( x=wind.sp, y= dew.point))+
  geom_point(mapping = aes(color=region))+
  geom_smooth(method = lm, mapping = aes(linetype = region))+
  facet_wrap(~region, nrow = 2)

```

