---
title: "Lab 04"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r install-libraries}
library(lubridate)
library(tidyverse)
library(data.table)
```


## Step 1. Read in the data
First download and then read in with data.table:fread()

```{r}
if (!file.exists("met_all.gz")){
  download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz", "met_all.gz", method="libcurl", timeout = 60)
}

met <- data.table::fread("met_all.gz")
```
## 2. Prepare the data

Remove temperatures less than -17C and change elev 9999 to missing value code.
```{r remove-low-temp}
met <- met[met$temp >= -17][elev == 9999.0, elev:= NA]
```

Generate a date variable using the functions as.Date() (hint: You will need the following to create a date paste(year, month, day, sep = "-")).

```{r create-ymd}
met <- met[, ymd := as.Date(paste(year, month, day, sep = "-"))]
```

Using the data.table::week function, keep the observations of the first week of the month.

```{r}
met[, table(week(ymd))]
met <- met[week(ymd)==31]
```

Compute the mean by station of the variables temp, rh, wind.sp, vis.dist, dew.point, lat, lon, and elev.

```{r check-max}
met[ , .(
  temp      = max(temp),
  rh        = max(rh, na.rm=T), 
  wind.sp   = max(wind.sp, na.rm=T), 
  vis.dist  = max(vis.dist, na.rm=T), 
  dew.point = max(dew.point, na.rm=T), 
  lat       = max(lat, na.rm=T), 
  lon       = max(lon, na.rm=T), 
  elev      = max(elev, na.rm=T)
)]
```
```{r}
met_avg <- met[ , .(
  temp      = mean(temp),
  rh        = mean(rh, na.rm=T), 
  wind.sp   = mean(wind.sp, na.rm=T), 
  vis.dist  = mean(vis.dist, na.rm=T), 
  dew.point = mean(dew.point, na.rm=T), 
  lat       = mean(lat, na.rm=T), 
  lon       = mean(lon, na.rm=T), 
  elev      = mean(elev, na.rm=T)
), by = "USAFID"]

```


Create a region variable for NW, SW, NE, SE based on lon = -98.00 and lat = 39.71 degrees
Create a categorical variable for elevation as in the lecture slides
## Step 2. Check the dimensions, headers, footers. How many columns, rows are there?


## Replace elevations with 9999 as NA.

```{r}
met[met$elev==9999.0] <- NA
met[, summary(elev)]
```
The weather station with the highest elevation is at `r max(met$elev, na.rm= TRUE)` meters.

```{r}
met <- met[temp>-40]
met2 <- met[order(temp)]
head(met2)
```

```{r}
met <- met[temp>-15][order(temp)]
```

## Step 5. Check the data against an external data source.


## Step 6. Calculate summary statistics

```{r}
met[elev==max(elev, na.rm=TRUE), summary(wind.sp)]
met[elev==max(elev, na.rm=TRUE), summary(temp)]
```
```{r}
met[elev==max(elev, na.rm=TRUE), .(
  temp_wind    = cor(temp, wind.sp, use= "complete"),
  temp_hour    = cor(temp, hour,    use ="complete"),
  wind.sp_day  = cor(wind.sp,day,   use ="complete"),
  wind.sp_hour = cor(wind.sp, hour, use ="complete")
)]
```



## Step 7. Exploratory graphs
```{r}
hist(met$elev, breaks=100)
hist(met$temp)
hist(met$wind.sp)
```

```{r}
library(leaflet)
library(dplyr)
library(tidyverse)
elev <- met[elev==max(elev)]
```

```{r}
#leaflet(elev) %>%
#  addProviderTiles('OpenStreetMap') %>% 
#  addCircles(lat=~lat,lng=~lon, opacity=1, fillOpacity=1, radius=100)
```