---
title: "Lab 03"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 1. Read in the data
First download and then read in with data.table:fread()

```{r}
if (!file.exists("met_all.gz")){
  download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz", "met_all.gz", method="libcurl", timeout = 60)
}

met <- data.table::fread("met_all.gz")
```

## Step 2. Check the dimensions, headers, footers. How many columns, rows are there?

```{r}
dim(met)
head(met)
tail(met)
```
## Step 3. Take a look at the variables.

```{r}
str(met)
```


## Step 4. Take a closer look

```{r}
table(met$year)
table(met$day)
table(met$hour)
```

```{r}
summary(met$temp)
summary(met$elev)
summary(met$wind.sp)
```

## Replace elevations with 9999 as NA.

```{r}
met[met$elev==9999.0] <- NA
met[, summary(elev)]
```
The weather station with the highest elevation is at `r max(met$elev, na.rm= TRUE)` meters.

```{r}
met <- met[temp>-40]
met2 <- met[order(temp)]
head(met2)
```

```{r}
met <- met[temp>-15][order(temp)]
```

## Step 5. Check the data against an external data source.


## Step 6. Calculate summary statistics

```{r}
met[elev==max(elev, na.rm=TRUE), summary(wind.sp)]
met[elev==max(elev, na.rm=TRUE), summary(temp)]
```
```{r}
met[elev==max(elev, na.rm=TRUE), .(
  temp_wind    = cor(temp, wind.sp, use= "complete"),
  temp_hour    = cor(temp, hour,    use ="complete"),
  wind.sp_day  = cor(wind.sp,day,   use ="complete"),
  wind.sp_hour = cor(wind.sp, hour, use ="complete")
)]
```



## Step 7. Exploratory graphs
```{r}
hist(met$elev, breaks=100)
hist(met$temp)
hist(met$wind.sp)
```

```{r}
library(leaflet)
library(dplyr)
library(tidyverse)
elev <- met[elev==max(elev)]
```

```{r}
leaflet(elev) %>%
  addProviderTiles('OpenStreetMap') %>% 
  addCircles(lat=~lat,lng=~lon, opacity=1, fillOpacity=1, radius=100)
```