---
title: "Lab 03"
output: html_document
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 1. Read in the data
First download and then read in with data.table:fread()

```{r}
if (!file.exists("met_all.gz")){
  download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/02_met/met_all.gz", "met_all.gz", method="libcurl", timeout = 60)
}

met <- data.table::fread("met_all.gz")
```

## Step 2. Check the dimensions, headers, footers. How many columns, rows are there?

```{r}
dim(met)
head(met)
tail(met)
```
## Step 3.

```{r}
str(met)
```


## Step 4. Take a closer look

```{r}
table(met$year)
table(met$day)
table(met$hour)
```

```{r}
summary(met$temp)
summary(met$elev)
summary(met$wind.sp)
```

##Replace elevations with 9999 as NA.

```{r}
met[met$elev==9999.0] <- NA
met[, summary(elev)]
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
