---
title: "Lab 06"
date: "`r Sys.Date()`"
output: github_document
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r install-libraries}
library(tidytext)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(forcats)
```


## Step 1. Read in the data
First download and then read in with read.csv()

```{r read-data, cache=TRUE}
if (!file.exists("mtsamples.csv")){
  download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/00_mtsamples/mtsamples.csv",
                "mtsamples.csv", method="libcurl", timeout = 60)
}

mts <- read.csv("mtsamples.csv")
str(mts)
mts <- as_tibble(mts)
mts
```

# Question 1: What specialties do we have?
We can use count() from dplyr to figure out how many records are there per speciality.

```{r medical-specialties}
specialties <- 
  mts %>%
  count(medical_specialty)

specialties %>% 
  arrange(desc(n)) %>%
knitr::kable()
```

These are `r nrow(specialties)` medical specialties.

```{r barplot-of-specialty-counts}
specialties %>%
  top_n(10) %>%
  ggplot( aes(x = n, y= fct_reorder(medical_specialty, n))) +
  geom_col()
```

The distribution is not at all uniform.

# Question 2
Tokenize the the words in the transcription column
Count the number of times each token appears
Visualize the top 20 most frequent words
Explain what we see from this result. Does it makes sense? What insights (if any) do we get?

```{r token-transcription, cache = TRUE}
mts %>%
  unnest_tokens(token, transcription) %>%
  count(token, sort = TRUE) %>%
  top_n(20, n) %>%
  ggplot(aes(n, fct_reorder(token, n))) +
  geom_col()
```

There are a lot of stopwords here, non-specific to medical text. We do see "patient", phew!

# Question 3. Redo analysis in Q2, and remove stopwords.
Redo visualization but remove stopwords before
Bonus points if you remove numbers as well

```{r, cache=TRUE}
mts %>%
  unnest_tokens(word, transcription) %>%
  count(word, sort = TRUE) %>%
  anti_join(stop_words, by = c("word")) %>%
  # use regular expression to filter out numbers
  filter( !grepl(pattern = "^[0-9]+$", x = word )) %>%
  top_n(20, n) %>%
  ggplot(aes(n, fct_reorder(word, n))) +
  geom_col()
```

